EDIpack 2.0: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-21 02_install.tex
GB wide pass on many parts of the text

The default target builds and install both the main library and the C-bindings.
Separate build targets for either are available. A recap message is printed at the end of the
CMake configuration step.

\subsubsection{Anaconda}
As an alternative we provide for both Linux and Mac Os systems
installation through Anaconda packages into a virtual
environment containing Python ($>3.10$).

>>>>

The default target builds and installs both the main library and the C-bindings.
Separate build targets for either are available. A recap message is printed at the end of the
CMake configuration step.

\subsubsection{Anaconda}
As an alternative, we provide for both Linux and macOS systems
installation through Anaconda packages into a virtual
environment containing Python ($>3.10$).

### 2025-05-21 03_edipack2.tex
GB wide pass on many parts of the text

The quantum states in the space $\FF$ are naturally represented in
the occupation number formalism of second quantization,
i.e., the Fock basis.
For a system of $N_\mathrm{s}$ electrons, each Fock state
is given as $\ket{p}\ket{\vec{n}}$, with
\begin{equation} \label{eq:FockState_|nupndw>}
    \ket{\vec{n}}=\ket{n_{1\up},\dots,n_{N_\mathrm{s}\up},n_{1\dw},\dots,n_{N_\mathrm{s}\dw}},
\end{equation}
where $p=1,\dots,N_\mathrm{ph}$ is the number of local phonons and
$n_{a\sigma}=0,1$ indicates the absence or 
presence of an electron with spin $\sigma$ at level $a$.

@@

\begin{itemize}
\item{} The {\bf normal} case conserves both the total occupation $N$
and the total magnetization $S_z$, or equivalently, the total number
of electrons with spin up $N_\up$ and down $N_\dw$.
Optionally, the symmetry
can be extended to act indipendently to each orbital and spin
component, i.e., $\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This
specific case is discussed extensively
Ref.~[\onlinecite{Amaricci2022}] and will not be further covered here.

@@

Alternatively, a list of two-body operators can be provided by the user in a properly
formatted text file, whose name is specified by the input variable {\tt
  UMATRIX\_FILE}. The implementation of the list of interaction terms is contained in
the class {\tt ED\_PARSE\_UMATRIX}.  
The file mush have the suffix {\tt \\.restart} and should be formatted
as follows: 
\begin{lstlisting}[style=mybash]
NORB BANDS
i1 j1 k1 l1 U_i1j1k1l1
i2 j2 k2 l2 U_i2j2k2l2
...
\end{lstlisting}
where {\tt NORB} is the numer of orbitals $N_\a$ and each line
defines a single two-body term of the form 
$U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}$. Note tha the
coefficient indices are swapped compared to the standard two-body
operator formulation.  
Empty lines and lines starting with {\tt \#, \%, !} are ignored.

@@

Each instance of {\tt sparse\_matrix\_csr} can be stored in either 
a serial manner (one copy per process) or distributed across multiple 
processes, with rows dynamically assigned to each process. 
Matrix elements are inserted using the 
{\tt sp\_insert\_element} procedure, which leverages the Fortran 
intrinsic {\tt move\_alloc} for faster execution compared to 
implicit reallocation, i.e., {\tt vec=[vec,new\_element]}.

@@

\subsubsection{Sparse map}\label{CodeSparseMap}
\fixme{GB: I don't like the minipage of the listing below.}
As discussed in \secu{sSecRDM}, the construction of a symmetry 
sector often requires associating each sector state $\ket{i}$ with 
disjoint information about the corresponding Fock state $\ket{I}$. 
In this context, the Fock state can be decomposed into bit chunks, 
e.g., $\ket{\vec{n}} = \ket{\vec{i}_\up\,\vec{b}_\up\,
\vec{i}_\dw\,\vec{b}_\dw}$, reflecting a natural grouping into impurity 
and bath components and where we highlight the overall ordering in
spin-up and spin-down parts (see \equ{eq:FockState_|nupndw>}).

@@

In both {\bf superc} and {\bf nonsu2} modes a single sparse map is
used.
Here, each Fock state $\ket{J}$ is associated with four integers $I_\up, B_\up, I_\dw,
B_\dw$ representing the bit decomposition of the state. 
These integers satisfy the relationship
$$
J =I_\up +  B_\up2^{N_\mathrm{imp}} + (I_\dw +  B_\dw2^{N_\mathrm{imp}})2^{N_\mathrm{s}}
$$
where $N_\mathrm{imp}$ is the number of impurity bits and 
$N_\mathrm{s}$ is the total number of spin components.
To obtain a contiguous memory pattern we define the \emph{key} for the sparse
map as $I=I_\up + I_\dw2^{N_\mathrm{imp}}$ and the corresponding
\emph{values} as the integers $B=B_\up + B_\dw2^{N_\mathrm{b}}$.
Note that these relations can be easily inverted to obtain $I_{\sigma=\up,\dw}$ from
$I$ and $B_{\sigma=\up,\dw}$ from $B$ in order to reconstruct the Fock
state $\ket{J}$.
The four integers satisfy the symmetry constraints:   
$\mathtt{popcnt}(I_\up + B_\up2^{N_\mathrm{imp}}) + \mathtt{popcnt}(I_\dw + B_\dw2^{N_\mathrm{imp}}) = N$
for the \textbf{nonsu2} case and
$\mathtt{popcnt}(I_\up + B_\up2^{N_\mathrm{imp}}) - \mathtt{popcnt}(I_\dw + B_\dw2^{N_\mathrm{imp}}) = S_z$
for the \textbf{superc} case.
This structure ensures an efficient and compact representation 
of the many-body Hilbert space, facilitating rapid state lookup 
during the diagonalization process.

@@

The bath topology, which defines the connectivity between the
$N_\mathrm{b}$ electronic levels assigned to the bath, is specified
through the input variable {\tt bath\_type}. This variable can take
one of four possible values: {\bf normal}, {\bf hybrid}, {\bf
  replica}, and {\bf general} (see Fig.~\ref{figBaths}).

\paragraph{{\tt bath\_type}={\bf normal}} In this configuration, the
bath consists  of $N_\mathrm{bath}$
electronic levels coupled to each of the impurity orbital, resulting
in a total  $N_\mathrm{b} = N_\a N_\mathrm{bath}$ of bath levels.

For {\tt ed\_mode}={\bf normal} the bath Hamiltonian is parametrized
as diagonal matrices in both the orbital and spin spaces:
$$
h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'} .
$$
where $\varepsilon_{\a\s}$ represents  the on-site energies for each orbital and spin.
If {\tt ed\_mode}={\bf superc},  anomalous
amplitudes connecting bath levels with opposite spins must also be
considered. In this case, the bath Hamiltonian gains 
an additional set of parameters, diagonal in orbital space but
off-diagonal in the spin one:
$$
h^p_{\a\b\s\s'}=\Delta^p_{\a}\delta_{\a\b}\delta_{\s \overline{\s}} .
$$
where $\Delta^p_\a$ represents the pairing amplitudes. 
This choice corresponds to a Nambu representation for each orbital component,
with the Hamiltonian matrices taking the form:
$$
\hat{h}^p_{\a\b} = \e_{\a}\delta_{\a\b}\tau_0 +
\Delta_{\a}\delta_{\a\b}\tau_x .
$$
where $\tau_{i=0,x,y,z}$ are Pauli matrices.

The hybridization amplitudes between the impurity and the bath levels
are similarly structured. For all values of {\tt ed\_mode} these
parameters are diagonal in both
spin and orbital space: $V^p_{\a\b\s\s'}=V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.
If {\tt ed\_mode}={\bf nonsu2} the an additional set including terms
describing spin-flip processes (as total magnetization is not
conserved) should be included $V^p_{\a\b\s\s'} =
W^p_{\a}\delta_{\a\b}\delta_{\s \overline{\s}}$.

\paragraph{{\tt bath\_type}={\bf hybrid}}
The bath is formed by $N_\mathrm{bath}$ levels coupled to all the
impurity orbitals, resulting in a total of 
$N_\mathrm{b}\equiv N_\mathrm{bath}$ electronic bath levels. 
The parametrization of the bath Hamiltonian is 
diagonal in both orbital and spin space, i.e.,
$$
h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'}.
$$

@@

If spin group is not conserved, i.e., for {\tt ed\_mode}={\bf nonsu2}, an additional
set of parameters must be specified to account for spin-flip processes:
$V^p_{\a\b\s\s'} = W^p_{\a\b}\delta_{\s \overline{\s}}$.

This hybrid topology of the bath captures the effects of locally
hybridized impurity orbitals, but typically requires a more
challenging optimization process, specially
with few available bath levels (see \secu{sSecFit}).

\paragraph{{\tt bath\_type}={\bf replica}/{\bf general}}
This configuration provides a more flexible parametrization of the 
bath. The core idea is to assign each bath element a structure 
that {\it replicates} the internal structure of the impurity, while 
maintaining diagonal coupling between bath elements and the impurity. 
This offloads the complexity of representing structured quantum 
impurities to the bath Hamiltonian rather than the hybridizations.

From a broader perspective, this can be formalized by considering  
a user-defined matrix basis
$\vec{\Gamma}=\{ \Gamma^\nu_{\a\b\s\s'} \}_{\nu=1,\dots,N_{sym}}$ in the (Nambu-)spin-orbital
space, and expressing the bath Hamiltonian as
$$
h^p = \sum_{\nu=1}^{N_{sym}} \lambda^p_\nu
\Gamma^\nu\equiv \vec{\lambda}^p\cdot \vec{\Gamma}
$$
where $\vec{\lambda}^p\in\RRR^{N_{sym}}$ is a vector of variational parameters. 
The choice of the matrix basis can be inspired by the internal structure
of the quantum impurity, i.e. $\hat{h}^0$, or determined
case-by-case by based on the physical properties of the system.
In this scheme, the total number of bath levels becomes: $N_\mathrm{b}
= N_{sym} N_\mathrm{bath}$.

For the {\bf replica} topology, the coupling between the impurity 
and each bath element remains diagonal in spin, orbital, and internal 
bath structure: $V^p_{\a\b\s\s'} = V^p$.
In the {\bf general} case, this is generalized to include explicit 
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.

@@

\paragraph{Bath Replica (General)}
The module {\tt ED\_BATH\_REPLICA} defines the class {\tt Hreplica} 
for the {\bf replica} and {\bf general} bath parametrizations. In 
addition to standard object construction, destruction, reading, and 
saving, this module provides matrix basis setup 
$\{ \Gamma^\nu \}_{\nu=1,\dots,N_{sym}}$ and variational parameter 
initialization through {\tt set\_Hreplica} (or {\tt set\_Hgeneral}). 
The actual bath Hamiltonian is constructed with the function 
{\tt build\_Hreplica} (or {\tt build\_Hgeneral}) via the relation
$
h^p = \vec{\lambda}^p \cdot \Gamma.
$

@@

% \begin{table}%[ht]
%   \label{TabBath}
% \begin{center}
% \begin{tabularx}{\linewidth}{ |X|X|X|X| } 
%   \hline
%   c  & {\bf normal} & {\bf superc} & {\bf  nonsu2} \\
%   \hline
%   {\bf normal} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_\a]$ & $\vec{\e}[N_\mathrm{b}],\vec{V}[N_\mathrm{b}]$ 
%                     & $\vec{\e}[N_\mathrm{b}], \vec{\Delta}[N_\mathrm{b}],\vec{V}[N_\mathrm{b}]$
%                                    &  $\vec{\e}[N_\mathrm{b}], \vec{V}[N_\mathrm{b}],\vec{W}[N_\mathrm{b}]$ \\
%   \hline
%   {\bf hybrid} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_\a]$ & $\vec{\e}[N_\mathrm{bath}],\vec{V}[N_\mathrm{b}]$ 
%                     & $\vec{\e}[N_\mathrm{bath}], \vec{\Delta}[N_\mathrm{bath}],\vec{V}[N_\mathrm{b}]$
%                                    &  $\vec{\e}[N_\mathrm{bath}], \vec{V}[N_\mathrm{b}],\vec{W}[N_\mathrm{b}]$ \\
%   \hline
%   {\bf replica} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_{sym}]$ & \multicolumn{3}{*{\hat{h} } \cr
%   \hline
%   {\bf general} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_{sym}]$ & \cline{2-4} & b&  c  \cr
%   \hline
% \end{tabularx}
% \end{center}
% \caption{Ciao}
% \end{table}

@@

The core computational task in any Krylov-based diagonalization
algorithm is the Matrix-Vector Product (MVP), i.e., the application
of the Hamiltonian to a given input vector:
\begin{equation}
H_\SS \ket{v} \to \ket{w}\qquad \ket{v},\ket{w}\in \SS_{\vec{Q}}.
\end{equation}
This operation easily accounts for more than 80\% of the overall
computational effort, making its optimization a critical aspect of
high-performance diagonalization algorithms. In \onlinecite{Amaricci2022},
the distributed-memory, massively parallel strategies employed in \NAME
to optimize this step are discussed in full detail. Here, we provide
a brief outline of the main ideas, following the notation introduced
in \onlinecite{Amaricci2022}, and focusing on the impact of the choice
of {\tt ed\_mode}.

@@

\subsection{Reduced impurity density matrix}\label{sSecRDM}
The latest version of \NAME introduces the calculation of the 
impurity Reduced Density Matrix (iRDM, $\rho^\mathrm{imp}$), 
extending the algorithm initially proposed in 
Ref.~\onlinecite{BellomiaPhD} to support the analysis of entanglement 
properties of quantum impurities for any value of  {\tt ed\_mode}.

For simplicity, this section focuses on the zero-temperature limit, 
assuming a non-degenerate ground state $\ket{\psi}$ present in the 
{\tt state\_list}. The generalization to the finite-temperature regime 
or degenerate ground states is straightforward, requiring only the 
replacement of the pure state average with the thermal or degenerate 
ensemble average $\ibra \rho^\mathrm{imp} \iket$, as defined in 
\equ{eq:thermal_average}.

A pure quantum state $\ket{\psi}$ belonging to a unique symmetry 
sector $\SS_{\vec{Q}}$ can be represented in the Fock basis as:
$
\ket{\psi} = \sum_I a_I \ket{I},
$
where the pure density matrix $\rho$ is then given by:
\begin{equation}
\rho = \ket{\psi}\bra{\psi} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
a^*_J a_I \ket{I}\bra{J} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
\rho_{IJ} \ket{I}\bra{J}.
\end{equation}

The iRDM is obtained by tracing out the bath degrees of freedom:
\begin{equation}
\rho^\mathrm{imp} = \Tr_\mathrm{bath}(\rho).
\end{equation}
However, the summations involved in this trace quickly become 
computationally prohibitive, as their range scales exponentially 
with the system size, i.e., $4^{N_\mathrm{s}}$.

To address this, we implemented a fast algorithm that exploits the 
block structure of the Fock space defined by the symmetry sectors and
the subsequent map sparsity. 
This approach significantly reduces the computational cost by limiting 
the summation to the sector size $D_\SS$ of the symmetry sector 
$\SS$ to which the ground state (or any eigenstate) belongs.

Next, we decompose the Fock basis into spin-dependent impurity and 
bath parts. Any Fock state $\ket{I}$ can be split as:
$
\ket{I} = \ket{i_\uparrow \, b_\uparrow} \otimes 
\ket{i_\downarrow \, b_\downarrow},
$
where the impurity $i_\sigma$ and bath $b_\sigma$ states span subspaces of dimension 
$D^\s_\mathrm{imp}$ and $D^\s_\mathrm{bath}$, respectively, for each 
spin $\s = \uparrow, \downarrow$.

To efficiently manage this decomposition, we introduce a 
{\tt sparse\_map} ${\cal P}$, as detailed in \secu{CodeSparseMap}. 
This structure stores, for each impurity configuration 
$i_\s = 1, \dots, D^\s_\mathrm{imp}$ (the \emph{key}), the 
corresponding bath configurations 
$b_\s \in [1, \dots, D^\s_\mathrm{bath}]$ (the \emph{values}) that 
satisfy the symmetry constraints over the sector quantum numbers
imposed by the {\tt ed\_mode} variable.
We denote the number of keys as ${\cal D}_{i\s}$ and the 
number of values as ${\cal D}_{b\s}$.
The fast summation algorithm for evaluating the iRDM differs 
significantly between the {\bf normal} and {\bf superc/nonsu2} modes, 
as outlined in the sections below.

\paragraph{{\bf normal}}
In this case, we take advantage of the Fock space factorization and 
sector spin-resolved decomposition, as discussed in detail in 
\secu{sSecQNs}. 
The global bitset can be split into spin-resolved 
impurity and bath components as:
\begin{equation}
\ket{\vec{n}} =
\ket{\vec{n}_\up}\otimes \ket{\vec{n}_\dw} =
\ket{\vec{i}_\up}\ket{\vec{b}_\up}\otimes \ket{\vec{i}_\dw}\ket{\vec{b}_\dw}.
\end{equation}
Here, creation (destruction) operators  $c^+_{p\sigma}$
($c^+_{p\sigma}$) at position $p$ and for a given spin $\sigma$
act only on the corresponding 
spin subspace, i.e., they commute with the opposite spin 
subspace $\ket{\vec{n}_{\bar{\s}}}$ without introducing any 
fermionic sign.

>>>>

The quantum states in the space $\FF$ are naturally represented in
the occupation number formalism of second quantization,
i.e.~the Fock basis.
For a system of $N_\mathrm{s}$ electrons, each Fock state
is given as $\ket{p}\ket{\vec{n}}$, with
\begin{equation} \label{eq:FockState_|nupndw>}
    \ket{\vec{n}}=\ket{n_{1\up},\dots,n_{N_\mathrm{s}\up},n_{1\dw},\dots,n_{N_\mathrm{s}\dw}},
\end{equation}
where $p=1,\dots,N_\mathrm{ph}$ is the number of local phonons and
$n_{a\sigma}=0,1$ indicates the absence or 
presence of an electron with spin $\sigma$ at level $a$.

@@

\begin{itemize}
\item{} The {\bf normal} case conserves both the total occupation $N$
and the total magnetization $S_z$, or equivalently, the total number
of electrons with spin up $N_\up$ and down $N_\dw$.
Optionally, the symmetry
can be extended to act indipendently to each orbital and spin
component, i.e.~$\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This
specific case is discussed extensively
Ref.~[\onlinecite{Amaricci2022}] and will not be further covered here.

@@

Alternatively, a list of two-body operators can be provided by the user in a properly
formatted text file, whose name is specified by the input variable {\tt
  UMATRIX\_FILE}. The implementation of the list of interaction terms is contained in
the module {\tt ED\_PARSE\_UMATRIX}.  
The file mush have the suffix {\tt \\.restart} and should be formatted
as follows: 
\begin{lstlisting}[style=mybash]
NORB BANDS
i1 j1 k1 l1 U_i1j1k1l1
i2 j2 k2 l2 U_i2j2k2l2
...
\end{lstlisting}
where {\tt NORB} is the numer of orbitals $N_\a$ and each line
defines a single two-body term of the form 
$U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}$. Note tha the
coefficient indices are swapped compared to the standard two-body
operator formulation.  
Empty lines and lines starting with {\tt \#, \%, !} are ignored.

@@

Each instance of {\tt sparse\_matrix\_csr} can be stored in either 
a serial manner (one copy per process) or distributed across multiple 
processes, with rows dynamically assigned to each process. 
Matrix elements are inserted using the 
{\tt sp\_insert\_element} procedure, which leverages the Fortran 
intrinsic {\tt move\_alloc} for faster execution compared to 
implicit reallocation, i.e.~{\tt vec=[vec,new\_element]}.

@@

\subsubsection{Sparse map}\label{CodeSparseMap}
As discussed in \secu{sSecRDM}, the construction of a symmetry 
sector often requires associating each sector state $\ket{i}$ with 
disjoint information about the corresponding Fock state $\ket{I}$. 
In this context, the Fock state can be decomposed into bit chunks, 
e.g., $\ket{\vec{n}} = \ket{\vec{i}_\up\,\vec{b}_\up\,
\vec{i}_\dw\,\vec{b}_\dw}$, reflecting a natural grouping into impurity 
and bath components and where we highlight the overall ordering in
spin-up and spin-down parts (see \equ{eq:FockState_|nupndw>}).

@@

In both {\bf superc} and {\bf nonsu2} modes a single sparse map is
used.
Here, each Fock state $\ket{J}$ is associated with four integers $I_\up, B_\up, I_\dw,
B_\dw$ representing the bit decomposition of the state. 
These integers satisfy the relationship
$$
J =I_\up +  B_\up2^{N_\mathrm{imp}} + (I_\dw +  B_\dw2^{N_\mathrm{imp}})2^{N_\mathrm{s}}
$$
where $N_\mathrm{imp}$ is the number of impurity bits and 
$N_\mathrm{s}$ is the total number of spin orbitals.
% To obtain a contiguous memory pattern %%> What? They are pointed lists...
We define the \emph{key} for the sparse
map as $I=I_\up + I_\dw2^{N_\mathrm{imp}}$ and the corresponding
\emph{values} as the integers $B=B_\up + B_\dw2^{N_\mathrm{b}}$.
Note that these relations can be easily inverted to obtain $I_{\sigma=\up,\dw}$ from
$I$ and $B_{\sigma=\up,\dw}$ from $B$ in order to reconstruct the Fock
state $\ket{J}$.
The four integers satisfy the symmetry constraints:   
$\mathtt{popcnt}(I_\up + B_\up2^{N_\mathrm{imp}}) + \mathtt{popcnt}(I_\dw + B_\dw2^{N_\mathrm{imp}}) = N$
for the \textbf{nonsu2} case and
$\mathtt{popcnt}(I_\up + B_\up2^{N_\mathrm{imp}}) - \mathtt{popcnt}(I_\dw + B_\dw2^{N_\mathrm{imp}}) = S_z$
for the \textbf{superc} case.
This structure ensures an efficient and compact representation 
of the many-body Hilbert space, enabling a rapid state lookup 
during the diagonalization process.

@@

The bath topology, which defines the connectivity between the
$N_\mathrm{b}$ electronic levels assigned to the bath, is specified
through the input variable {\tt bath\_type}. This variable can take
one of four possible values: {\bf normal}, {\bf hybrid}, {\bf
  replica}, and {\bf general} (see Fig.~\ref{figBaths}).
  \fixme{This figure looks quite annoying to generate, especially to cover
  all possible \texttt{ed\_mode} values. I'd pass...}

\paragraph{{\tt bath\_type}={\bf normal}} In this configuration, the
bath consists of $N_\mathrm{bath}$
electronic sites coupled to each of the impurity orbital, resulting
in a total  $N_\mathrm{b} = N_\sigma N_\a N_\mathrm{bath}$ of bath levels.
% Ocio!

For {\tt ed\_mode}={\bf normal} the bath Hamiltonian is parametrized
as diagonal matrices in both the orbital and spin spaces:
$$
h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'} .
$$
where $\varepsilon_{\a\s}$ represents  the on-site energies for each orbital and spin.

If {\tt ed\_mode}={\bf superc},  anomalous
amplitudes connecting bath levels with opposite spins must also be
considered. In this case, the bath Hamiltonian gains 
an additional set of parameters, diagonal in orbital space but
off-diagonal in the spin one:
$$
h^p_{\a\b\s\s'}=\Delta^p_{\a}\delta_{\a\b}\delta_{\s \overline{\s}} .
$$
where $\Delta^p_\a$ represents the pairing amplitudes. 
This choice corresponds to a Nambu representation for each orbital component,
with the Hamiltonian matrices taking the form:
$$
\hat{h}^p_{\a\b} = \e_{\a}\delta_{\a\b}\tau_0 +
\Delta_{\a}\delta_{\a\b}\tau_x .
$$
where $\tau_{i=0,x,y,z}$ are Pauli matrices.

The hybridization amplitudes between the impurity and the bath levels
are similarly structured. For all values of {\tt ed\_mode} these
parameters are diagonal in both
spin and orbital space: $V^p_{\a\b\s\s'}=V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.

If {\tt ed\_mode}={\bf nonsu2} the an additional set including terms
describing spin-flip processes (as total magnetization is not
conserved) should be included $V^p_{\a\b\s\s'} =
W^p_{\a}\delta_{\a\b}\delta_{\s \overline{\s}}$.

\paragraph{{\tt bath\_type}={\bf hybrid}}
The bath is formed by $N_\mathrm{bath}$ sites coupled to all the
impurity orbitals, resulting in a total of % Ocio!
$N_\mathrm{b}\equiv N_\sigma N_\alpha N_\mathrm{bath}$ electronic bath levels. 
The parametrization of the bath Hamiltonian is 
diagonal in both orbital and spin space, i.e.,
$$
h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'}.
$$

@@

If full spin symmetry is not preserved, i.e.~for {\tt ed\_mode}={\bf nonsu2}, an additional
set of parameters must be specified to account for spin-flip processes:
$V^p_{\a\b\s\s'} = W^p_{\a\b}\delta_{\s \overline{\s}}$.

This hybrid topology of the bath captures the effects of locally
hybridized impurity orbitals, but typically requires a more
challenging optimization process, especially
with few available bath levels (see \secu{sSecFit}).

\paragraph{{\tt bath\_type}={\bf replica}/{\bf general}}
This configuration provides a more flexible parametrization of the 
bath. The original core idea \cite{Capone2004PRB,Civelli2006}
is to assign each bath element a structure 
that {\it replicates} the internal structure of the impurity, while 
maintaining diagonal coupling between bath elements and the impurity. 
This offloads the complexity of representing structured quantum 
impurities to the bath Hamiltonian rather than the hybridizations.

From a broader perspective, this can be formalized by considering  
a user-defined matrix basis
$\vec{\Gamma}=\{ \Gamma^\nu_{\a\b\s\s'} \}_{\nu=1,\dots,N_\mathrm{sym}}$ in the (Nambu-)spin-orbital
space, and expressing the bath Hamiltonian as
$$
h^p = \sum_{\nu=1}^{N_\mathrm{sym}} \lambda^p_\nu
\Gamma^\nu\equiv \vec{\lambda}^p\cdot \vec{\Gamma}
$$
where $\vec{\lambda}^p\in\RRR^{N_\mathrm{sym}}$ is a vector of variational parameters. 
The choice of the matrix basis can be inspired by the internal structure
of the quantum impurity, i.e. $\hat{h}^0$, or determined
case-by-case, depending on the desired properties of the solution.
In this scheme, the total number of bath levels becomes $N_\mathrm{b}
= N_\mathrm{imp} N_\mathrm{bath}$, while the number of independent bath
parameters to optimize is (excluding the hybridization amplitudes)
$N_\mathrm{params} = N_\mathrm{sym} N_\mathrm{bath}$.
% Ocio!

For the {\bf replica} topology, the coupling between the impurity 
and each bath element is independent on spin, orbital, and internal 
bath structure: $V^p_{\a\b\s\s'} = V^p$.
In the {\bf general} case, this is generalized to include explicit 
diagonal
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.

@@

\paragraph{Bath Replica (General)}
The module {\tt ED\_BATH\_REPLICA} defines the class {\tt Hreplica} 
for the {\bf replica} and {\bf general} bath parametrizations. In 
addition to standard object construction, destruction, reading, and 
saving, this module provides matrix basis setup 
$\{ \Gamma^\nu \}_{\nu=1,\dots,N_\mathrm{sym}}$ and variational parameter 
initialization through {\tt set\_Hreplica} (or {\tt set\_Hgeneral}). 
The actual bath Hamiltonian is constructed with the function 
{\tt build\_Hreplica} (or {\tt build\_Hgeneral}) via the relation
$
h^p = \vec{\lambda}^p \cdot \Gamma.
$

@@

% \begin{table}%[ht]
%   \label{TabBath}
% \begin{center}
% \begin{tabularx}{\linewidth}{ |X|X|X|X| } 
%   \hline
%   c  & {\bf normal} & {\bf superc} & {\bf  nonsu2} \\
%   \hline
%   {\bf normal} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_\a]$ & $\vec{\e}[N_\mathrm{b}],\vec{V}[N_\mathrm{b}]$ 
%                     & $\vec{\e}[N_\mathrm{b}], \vec{\Delta}[N_\mathrm{b}],\vec{V}[N_\mathrm{b}]$
%                                    &  $\vec{\e}[N_\mathrm{b}], \vec{V}[N_\mathrm{b}],\vec{W}[N_\mathrm{b}]$ \\
%   \hline
%   {\bf hybrid} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_\a]$ & $\vec{\e}[N_\mathrm{bath}],\vec{V}[N_\mathrm{b}]$ 
%                     & $\vec{\e}[N_\mathrm{bath}], \vec{\Delta}[N_\mathrm{bath}],\vec{V}[N_\mathrm{b}]$
%                                    &  $\vec{\e}[N_\mathrm{bath}], \vec{V}[N_\mathrm{b}],\vec{W}[N_\mathrm{b}]$ \\
%   \hline
%   {\bf replica} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_\mathrm{sym}]$ & \multicolumn{3}{*{\hat{h} } \cr
%   \hline
%   {\bf general} $[N_\mathrm{b}\!=\!N_\mathrm{bath}N_\mathrm{sym}]$ & \cline{2-4} & b&  c  \cr
%   \hline
% \end{tabularx}
% \end{center}
% \caption{Ciao}
% \end{table}

@@

The core computational task in any Krylov-based diagonalization
algorithm is the Matrix-Vector Product (MVP), i.e.~the application
of the Hamiltonian to a given input vector:
\begin{equation}
H_\SS \ket{v} \to \ket{w}\qquad \ket{v},\ket{w}\in \SS_{\vec{Q}}.
\end{equation}
This operation easily accounts for more than 80\% of the overall
computational effort, making its optimization a critical aspect of
high-performance diagonalization algorithms. In \onlinecite{Amaricci2022},
the distributed-memory, massively parallel strategies employed in \NAME
to optimize this step are discussed in full detail. Here, we provide
a brief outline of the main ideas, following the notation introduced
in \onlinecite{Amaricci2022}, and focusing on the impact of the choice
of {\tt ed\_mode}.

@@

\subsection{Reduced impurity density matrix}\label{sSecRDM}
The latest version of \NAME introduces the calculation of the 
impurity Reduced Density Matrix (iRDM, $\rho^\mathrm{imp}$), 
extending the algorithm initially proposed in 
Ref.~\onlinecite{BellomiaPhD}, to support the analysis of entanglement 
properties of quantum impurities for any value of  {\tt ed\_mode}.

For simplicity, this section focuses on the zero-temperature limit, 
assuming a non-degenerate ground state $\ket{\psi}$ present in the 
{\tt state\_list}. The generalization to the finite-temperature regime 
or degenerate ground states is straightforward, requiring only the 
% replacement of the pure state average %%> What is a pure state average?
introduction of a thermal or degenerate 
ensemble average $\ibra \rho^\mathrm{imp} \iket$, as defined in 
\equ{eq:thermal_average}.

A pure quantum state $\ket{\psi}$ belonging to a unique symmetry 
sector $\SS_{\vec{Q}}$ can be represented in the Fock basis as:
$
\ket{\psi} = \sum_I a_I \ket{I},
$
where the corresponding pure density matrix $\rho$ is given by:
\begin{equation}
\rho = \ket{\psi}\bra{\psi} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
a^*_J a_I \ket{I}\bra{J} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
\rho_{IJ} \ket{I}\bra{J}.
\end{equation}

The iRDM is obtained by tracing out the bath degrees of freedom:
\begin{equation}
\rho^\mathrm{imp} = \Tr_\mathrm{bath}(\rho).
\end{equation}
However, the memory footprint of $\rho$, as well as the CPU footprint 
of the summations involved in this trace scales exponentially 
with the system size, i.e.~$4^{N_\mathrm{s}}$, quickly becoming prohibitive.

To overcome the issue, we implement a fast algorithm that exploits the 
block structure of the Fock space defined by the symmetry sectors and
the subsequent map sparsity, to perform the trace \emph{on-the-flight}. 
This approach significantly reduces the computational cost by limiting 
the summation to the sector size $D_\SS$ of the symmetry sector 
$\SS$ to which the ground state (or any eigenstate) belongs, and avoiding
the storage of the large matrix associated with $\rho$.

%%% ATTENZIONE: this is true only for the normal case!
%
%Next, we decompose the Fock basis into spin-dependent impurity and 
%bath parts. Any Fock state $\ket{I}$ can be split as:
%$
%\ket{I} = \ket{i_\uparrow \, b_\uparrow} \otimes 
%\ket{i_\downarrow \, b_\downarrow},
%$
%where the impurity $i_\sigma$ and bath $b_\sigma$ states span subspaces of dimension 
%$D^\s_\mathrm{imp}$ and $D^\s_\mathrm{bath}$, respectively, for each 
%spin $\s = \uparrow, \downarrow$.
%
%%% I WOULD SAY THAT WE DON'T NEED TO REPEAT OURSELVES< THE SECTOR
%%% SECTION IS THERE FOR A REASON. I try to edit below:
To efficiently manage this decomposition, we introduce a 
{\tt sparse\_map} ${\cal P}$, as detailed in \secu{CodeSparseMap}. 
This structure stores, for each impurity configuration 
(the \emph{key}), the 
corresponding bath configurations 
(the \emph{values}) that 
satisfy the symmetry constraints over the sector quantum numbers
imposed by the {\tt ed\_mode} variable.
We denote the number of (spin-dependent) keys and values respectively
as ${\cal D}_{i\s}$ and ${\cal D}_{b\s}$.
The fast summation algorithm for evaluating the iRDM differs 
significantly between the {\bf normal} and {\bf superc/nonsu2} modes, 
as outlined in the sections below.

\paragraph{{\bf normal}}
In this case, we take advantage of spin-resolved sector decomposition, 
as discussed in detail in 
\secu{sSecQNs}. 
The global bitset can be split into spin-resolved 
impurity and bath components as:
\begin{equation}
\ket{\vec{n}} =
\ket{\vec{n}_\up}\otimes \ket{\vec{n}_\dw} =
\ket{\vec{i}_\up}\ket{\vec{b}_\up}\otimes \ket{\vec{i}_\dw}\ket{\vec{b}_\dw}.
\end{equation}
Here, creation (destruction) operators  $c^+_{p\sigma}$
($c^+_{p\sigma}$) at position $p$ and for a given spin $\sigma$
act only on the corresponding 
spin subspace, i.e.~they commute with the opposite spin 
subspace $\ket{\vec{n}_{\bar{\s}}}$ without introducing any 
fermionic sign.
